Value Ordering, Discrepancies, and Restarts for Subgraph Algorithms

### 2018-07-26 paper/paper.tex
A bit of editoring

                \nl \lForEach{$c \in C$}{$\textnormal{post nogood}~B \wedge (v \mapsto c) \Rightarrow \bot$\tikzmark{restart5r}}

@@

    \nl \While{$(D_v, v') \gets ($\textnormal{a unit domain from}~$D, \textnormal{the single value
    in that domain})$}{

@@

Other studies use a random selection of 200 of each of the instances from the ``meshes'' and
``images'' families because some earlier solvers find many of these instances extremely hard. We
would like to have a larger number of satisfiable instances in our test set, and so we include all
pattern / target pairs. This gives a total of 14,621 instances (rather than the original 5,725). At
least \input{gen-n-sat.tex}\unskip\ of these instances are known to be satisfiable, and at least
\input{gen-n-unsat.tex}\unskip\ are unsatisfiable.

@@

\item[Biased] We branch by selecting a vertex $v'$ from the chosen domain $D_v$ with
probability \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \text{.} \] This is a new
heuristic, which continues to prefer vertices of high degree, but will give the same chance of
being selected next to two vertices of equal degree.  It also introduces an element of randomness,
which is needed to make restarts have an effect. \end{description}

The performance of the degree and biased heuristics are very similar: the
biased heuristic has slightly worse performance below a 400s timeout, and effectively equal
performance above it. This shows that we can introduce an element of randomness into the degree
value-ordering heuristic without adversely affecting its performance \emph{in aggregate}.
\Cref{figure:scatter-heuristics} gives a more detailed look at these results. This scatter plot
shows that despite the aggregate performance being very similar, on a case by case basis, the two
heuristics can make a large difference to the performance of satisfiable instances.

>>>>

                \nl \ForEach{$c \in C$}{\nl$\textnormal{post nogood}~B \wedge (v \mapsto c) \Rightarrow \bot$\tikzmark{restart5r}}

@@

    \nl \While{$D_v \gets $\textnormal{a unit domain from}~$D$}{
        \nl $v' \gets \textnormal{the single value in}~D_v$ \;

@@

\noindent
Other studies use a random selection of 200 of each of the instances from the ``meshes'' and
``images'' families because some earlier solvers find many of these instances extremely hard. We
would like to have a larger number of satisfiable instances in our test set, and so we include all
pattern / target pairs. This gives a total of 14,621 instances (rather than the original 5,725). At
least \input{gen-n-sat.tex}\unskip\ of these instances are known to be satisfiable, and at least
\input{gen-n-unsat.tex}\unskip\ are unsatisfiable.

@@

\item[Biased] We branch by selecting a vertex $v'$ from the chosen domain $D_v$ with
probability \[ p(v') = \frac{2^{\deg(v')}}{\sum_{w \in D_v}{2^{\deg(w)}}} \text{.} \] This is a new
heuristic, which continues to prefer vertices of high degree, but will give the same chance of
being selected next to two vertices of equal degree.  It also introduces an element of randomness,
which will be needed to make restarts have an effect. \end{description}

\noindent
The performance of the degree and biased heuristics are very similar: the
biased heuristic has slightly worse performance below a 400s timeout, and effectively equal
performance above it. This shows that we can introduce an element of randomness into the degree
value-ordering heuristic without adversely affecting its performance \emph{in aggregate}.
\Cref{figure:scatter-heuristics} gives a detailed look at these results. This scatter plot
shows that despite the aggregate performance being similar, on a case by case basis, the two
heuristics can make a large difference to the performance of satisfiable instances.
